Discrete Quantum Theories and Computing

### 2018-03-13 dissertation.tex
dissertation.lyx "Chapter 2 Conventional Quantum Theory" Updating...

Entanglement may be regarded as one of the main characteristics distinguishing
quantum from classical mechanics. Entanglement involves quantum correlations
such that the measurement outcomes in one subsystem are related to
the measurement outcomes in another one. Within the standard framework,
given a quantum system composed of $n$ qubit subsystems, a pure state
of the total system $\ket{\Psi}$ is said to be entangled if it cannot
be written as a product of states of each subsystem. That is, a state
$\ket{\Psi}$ is entangled if $\ket{\Psi}\ne\ket{\psi_{1}}\otimes\cdots\otimes\ket{\psi_{j}}\otimes\cdots\otimes\ket{\psi_{n}}$,
where $\ket{\psi_{j}}$ refers to an arbitrary state of the $j$-th
qubit, and $\otimes$ represents the tensor product. This is equivalent
to saying that if one calculates the reduced density operator $\rho_{j}$
of the $j$-th subsystem by tracing out all the other subsystems,
$\rho_{j}=\Tr_{\{1,\cdots,j-1,j+1,\cdots,n\}}\left(\rho\right)$,
with $j=1,\cdots,n$ and $\rho=\op{\Psi}{\Psi}$, the normalized state~$\ket{\Psi}$
is entangled if and only if at least one subsystem state is \textit{mixed};
i.e., $\Tr_{j}\left(\rho_{j}^{2}\right)<1$. For example, consider
\begin{equation}
\rho_{j}=\frac{1}{2}\left(\sigma_{0}+\sum\limits _{\mu=x,y,z}\expval{\sigma_{\mu}^{j}}\sigma_{\mu}^{j}\right)\,,
\end{equation}
where $\sigma_{\mu}^{j}$, $\mu=x,y,z$, are the Pauli operators acting
on the $j$-th spin~\cite{544199}, 
\begin{equation}
\sigma_{\mu}^{j}=\overbrace{\sigma_{0}\otimes\cdots\otimes\sigma_{0}\otimes\underbrace{\sigma_{\mu}}_{j^{th}\ \mbox{factor}}\otimes\sigma_{0}\otimes\cdots\otimes\sigma_{0}}^{n\ \mbox{factors}}\,,\label{pauli3}
\end{equation}
with 
\begin{equation}
\begin{aligned}\sigma_{0} & =\left(\begin{array}{cc}
1 & 0\\
0 & 1
\end{array}\right)\,, & \sigma_{x} & =\left(\begin{array}{cc}
0 & 1\\
1 & 0
\end{array}\right)\,, & \sigma_{y} & =\left(\begin{array}{cc}
0 & -i\\
i & 0
\end{array}\right)\,, & \sigma_{z} & =\left(\begin{array}{cc}
1 & 0\\
0 & -1
\end{array}\right)\,,\end{aligned}
\end{equation}
and $\expval{\sigma_{\mu}^{j}}=\melem{\Psi}{\sigma_{\mu}^{j}}{\Psi}$
denotes the corresponding expectation value. The vectors 
\begin{equation}
{\bf X}_{j}=\left(\expval{\sigma_{x}^{j}},\expval{\sigma_{y}^{j}},\expval{\sigma_{y}^{j}}\right)\in\mathbb{R}^{3}
\end{equation}
allow a geometric representation of each reduced state in $\mathbb{R}^{3}$,
satisfying $0\le\left\Vert {\bf X}_{j}\right\Vert \le1$. Since $\Tr_{j}\left(\rho_{j}^{2}\right)=\frac{1}{2}\left(1+\left\Vert {\bf X}_{j}\right\Vert ^{2}\right)$,
the state $\ket{\Psi}$ is entangled if $\left\Vert {\bf X}_{j}\right\Vert <1$
for at least one $j$, represented by a point \emph{inside} the corresponding
local Bloch sphere. One may therefore consider $\ket{\Psi}$ to be
maximally entangled if $\left\Vert {\bf X}_{j}\right\Vert =0$ for
all $j$. On the other hand, the state $\ket{\Psi}$ is unentangled
(i.e., a product state) if $\left\Vert {\bf X}_{j}\right\Vert =1$
for all $j$, corresponding to points lying on the surface of the
Bloch sphere.

@@

\todo{Explain enough background knowledge to support the discussion
about contextuality in the end of Sec.~\ref{modalquantum}.}

>>>>

Entanglement may be regarded as one of the main characteristics distinguishing
quantum from classical mechanics. Entanglement involves quantum correlations
such that the measurement outcomes in one subsystem are related to
the measurement outcomes in another one. Within the standard framework,
given a quantum system composed of $n$ qubit subsystems, a pure state
of the total system $\ket{\Psi}$ is said to be entangled if it cannot
be written as a product of states of each subsystem. That is, a state
$\ket{\Psi}$ is entangled if $\ket{\Psi}\ne\ket{\psi_{1}}\otimes\cdots\otimes\ket{\psi_{j}}\otimes\cdots\otimes\ket{\psi_{n}}$,
where $\ket{\psi_{j}}$ refers to an arbitrary state of the $j$-th
qubit, and $\otimes$ represents the tensor product. This is equivalent
to saying that if one calculates the reduced density operator $\rho_{j}$
of the $j$-th subsystem by tracing out all the other subsystems,
$\rho_{j}=\Tr_{\{1,\cdots,j-1,j+1,\cdots,n\}}\left(\rho\right)$,
with $j=1,\cdots,n$ and $\rho=\op{\Psi}{\Psi}$, the normalized state~$\ket{\Psi}$
is entangled if and only if at least one subsystem state is \textit{mixed};
i.e., $\Tr_{j}\left(\rho_{j}^{2}\right)<1$. For example, consider
\begin{equation}
\rho_{j}=\frac{1}{2}\left(\sigma_{0}+\sum\limits _{\mu=x,y,z}\expval{\sigma_{\mu}^{j}}\sigma_{\mu}^{j}\right)\,,
\end{equation}
where $\sigma_{\mu}^{j}$, $\mu=x,y,z$, are the Pauli operators acting
on the $j$-th spin~\cite{544199}, 
\begin{equation}
\sigma_{\mu}^{j}=\overbrace{\sigma_{0}\otimes\cdots\otimes\sigma_{0}\otimes\underbrace{\sigma_{\mu}}_{j^{th}\ \mbox{factor}}\otimes\sigma_{0}\otimes\cdots\otimes\sigma_{0}}^{n\ \mbox{factors}}\,,\label{pauli3}
\end{equation}
with 
\begin{equation}
\begin{aligned}\sigma_{0} & =\left(\begin{array}{cc}
1 & 0\\
0 & 1
\end{array}\right)\,, & \sigma_{x} & =\left(\begin{array}{cc}
0 & 1\\
1 & 0
\end{array}\right)\,, & \sigma_{y} & =\left(\begin{array}{cc}
0 & -\rmi\\
\rmi & 0
\end{array}\right)\,, & \sigma_{z} & =\left(\begin{array}{cc}
1 & 0\\
0 & -1
\end{array}\right)\,,\end{aligned}
\end{equation}
and $\expval{\sigma_{\mu}^{j}}=\melem{\Psi}{\sigma_{\mu}^{j}}{\Psi}$
denotes the corresponding expectation value. The vectors 
\begin{equation}
{\bf X}_{j}=\left(\expval{\sigma_{x}^{j}},\expval{\sigma_{y}^{j}},\expval{\sigma_{y}^{j}}\right)\in\mathbb{R}^{3}
\end{equation}
allow a geometric representation of each reduced state in $\mathbb{R}^{3}$,
satisfying $0\le\left\Vert {\bf X}_{j}\right\Vert \le1$. Since $\Tr_{j}\left(\rho_{j}^{2}\right)=\frac{1}{2}\left(1+\left\Vert {\bf X}_{j}\right\Vert ^{2}\right)$,
the state $\ket{\Psi}$ is entangled if $\left\Vert {\bf X}_{j}\right\Vert <1$
for at least one $j$, represented by a point \emph{inside} the corresponding
local Bloch sphere. One may therefore consider $\ket{\Psi}$ to be
maximally entangled if $\left\Vert {\bf X}_{j}\right\Vert =0$ for
all $j$. On the other hand, the state $\ket{\Psi}$ is unentangled
(i.e., a product state) if $\left\Vert {\bf X}_{j}\right\Vert =1$
for all $j$, corresponding to points lying on the surface of the
Bloch sphere.

@@

\todo{Explain enough background knowledge to support the discussion
about contextuality in the end of Sec.~\ref{modalquantum}. I have
typed some literature review in writing course, but I need to decide
whether it could be used or not before finish Sec.~\ref{sec:fuzzy}.}
